---
title: "Case Study 2- Attrition Data Analysis"
author: "Feby Thomas Cheruvathoor"
date: "12.3.20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction


# Data Import

The following code imports the attrition data into R Studio, from here the data will be cleaned, transformed and modeled to answer various analysis questions and provide insight and key takeaways.

```{r}
library (readr)


attrition_urlfile="https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2-data.csv"
no_attrition_urlfile="https://raw.githubusercontent.com/BivinSadler/MSDS_6306_Doing-Data-Science/Master/Unit%2014%20and%2015%20Case%20Study%202/CaseStudy2CompSet%20No%20Attrition.csv"

attrition_table <-read_csv(url(attrition_urlfile))

no_attrition_table <-read_csv(url(no_attrition_urlfile))

```


# Data Cleanup and remover redudant data 

Following code drops the columns which have stale data points - ID,EmployeeCount,Over18,StandardHours


```{r message=FALSE, warning=FALSE}
library(tidyverse)

attrition_table <- subset(attrition_table, select =-c(ID,EmployeeCount,Over18,StandardHours))

attrition_table <- attrition_table %>% select(Attrition,everything())

```


Bivariate EDA Analysis to explore the independednt variables with target(Attrition) variable 
```{r message=FALSE, warning=FALSE}
library(gt)
library(RColorBrewer)
library(ggplot2)
library(gridExtra)


p1 <- attrition_table %>%
  ggplot(aes(x = Age, fill = Attrition)) +      scale_fill_brewer(palette="Set1") + geom_density(alpha = 0.5) + ggtitle("Age") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p2 <- attrition_table %>%
  ggplot(aes(x = YearsInCurrentRole, fill = Attrition)) +   scale_fill_brewer(palette="Set1")+ geom_density(alpha = 0.5) + ggtitle("Years In Current Role")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p3 <- attrition_table %>%
  ggplot(aes(x = NumCompaniesWorked, fill = Attrition)) +scale_fill_brewer(palette="Set1")+  geom_density(alpha = 0.5) + ggtitle("Number of Companies")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p4 <- attrition_table %>%
  ggplot(aes(x = TotalWorkingYears, fill = Attrition)) + scale_fill_brewer(palette="Set1")+ geom_density(alpha = 0.5) + ggtitle("Total Working Years")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p5 <- attrition_table %>%
  ggplot(aes(x = JobLevel, fill = Attrition)) + scale_fill_brewer(palette="Set1")+ geom_density(alpha = 0.5) + ggtitle("Job Level")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

p6 <- attrition_table %>%
  ggplot(aes(x = MonthlyIncome, fill = Attrition)) + scale_fill_brewer(palette="Set1")+ geom_density(alpha = 0.5) + ggtitle("Monthly Income")  + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())

grid.arrange(p1, p2, p3, p4,p5,p6, nrow = 2, ncol = 3)


```

Additional Attrition Exploratory Analysis 

```{r message=FALSE, warning=FALSE}

p1 <- attrition_table %>%
  ggplot(aes(x = PercentSalaryHike, fill = Attrition)) + scale_fill_brewer(palette="Set1")+geom_density(alpha = 0.5) + ggtitle("Percentage Salary Hike") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


p2 <- attrition_table %>%
  ggplot(aes(x = YearsAtCompany, fill = Attrition)) + scale_fill_brewer(palette="Set1")+geom_density(alpha = 0.5) + ggtitle("Years At Company") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())



p3 <- attrition_table %>%
  ggplot(aes(x = YearsSinceLastPromotion, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years Since Last Promotion") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


p4 <- attrition_table %>%
  ggplot(aes(x = YearsWithCurrManager, fill = Attrition)) + geom_density(alpha = 0.5) + ggtitle("Years With Current Manager") + theme(plot.title = element_text(size =10),axis.text.x = element_text(size =7,angle = 45, hjust = 1),axis.title.x=element_blank())


grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)

```


###### Change  the class variables and the levels  for the following vaiables


```{r message=FALSE, warning=FALSE}

attrition_table$BusinessTravel <- factor(attrition_table$BusinessTravel, levels =  c('Non-Travel','Travel_Rarely','Travel_Frequently'),
                                         label = c('Non-Travel','Travel_Rarely','Travel_Frequently'))

attrition_table$Department <- factor(attrition_table$Department, levels =  c('Sales','Research & Development','Human Resources'),
                                     label = c('Sales','Research & Development','Human Resources'))

attrition_table$EducationField <- factor(attrition_table$EducationField, levels =  c('Life Sciences','Medical','Technical Degree','Other','Marketing',
                                                                                     'Human Resources'),label = c('Life Sciences','Medical','Technical Degree','Other','Marketing','Human Resources'))

attrition_table$Gender <- factor(attrition_table$Gender, levels =  c('Male','Female'),label = c('Male','Female'))

attrition_table$JobRole <- factor(attrition_table$JobRole, levels =  c('Sales Executive','Sales Representative','Research Director','Research Scientist','Manufacturing Director',
                                                                       'Healthcare Representative','Manager','Human Resources','Laboratory Technician'),label = c('Sales Executive','Sales Representative','Research Director','Research Scientist',
                                                                                                                                                                  'Manufacturing Director','Healthcare Representative','Manager','Human Resources','Laboratory Technician'))

attrition_table$MaritalStatus <- factor(attrition_table$MaritalStatus,levels =  c('Single','Divorced','Married'),label = c('Single','Divorced','Married'))

attrition_table$OverTime <- factor(attrition_table$OverTime, levels =  c('No','Yes'),label = c('No','Yes'))

attrition_table[sapply(attrition_table,is.factor)]  <- data.matrix(attrition_table[sapply(attrition_table,is.factor)])

attrition_table_org <- attrition_table

attrition_table$Attrition  <- factor(attrition_table$Attrition, levels =  c('No','Yes'),label = c('No','Yes'))

attrition_table[sapply(attrition_table,is.factor)]  <- data.matrix(attrition_table[sapply(attrition_table,is.factor)])

```

# Show the correlation of variables in the attrition table 


```{r message=FALSE, warning=FALSE}

library(corrplot)

attrition_table_cor <- attrition_table

attrition_table_cor.cor = cor(attrition_table_cor)

corrplot(attrition_table_cor.cor,method ="number", type="lower", lower.col = "black",tl.col="black", number.cex = .7)

```

#KNN Leave One Out 


```{r message=FALSE, warning=FALSE}
library(class)
library(caret)

iterations = 100
numks = 30

nCol  = ncol(attrition_table_org)

data_norm <- function(x) { (x -min(x))/(max(x)-min(x))   }

attrition_table_norm <- as.data.frame(lapply(attrition_table_org[2:nCol],data_norm)) # Omit the Attrition filed


attrition_table_train <- attrition_table_norm[1:696,]
attrition_table_test  <- attrition_table_norm[697:870,]

attrition_table_train_labels <- attrition_table_org[1:696,1]
attrition_table_test_labels  <- attrition_table_org[697:870,1]


masterAcc_l = matrix(nrow=iterations, ncol = numks)
masterSens_l = matrix(nrow=iterations, ncol = numks)
masterSpef_l = matrix(nrow=iterations, ncol = numks)

for(j in 1:iterations)
{
  accs_l = data.frame(accuracy = numeric(30),k=numeric(30))
  
  for(i in 1:30)
  {
    
    classifications_l = knn.cv(attrition_table_train,attrition_table_train_labels$Attrition,prob=TRUE,k=i)
    
    CM_l = confusionMatrix(table(attrition_table_train_labels$Attrition,classifications_l)) 
    masterAcc_l[j,i] = CM_l$overall[1]
    masterSpef_l[j,i] = CM_l$byClass[2]
    masterSens_l[j,i] = CM_l$byClass[1]
    
  }
}

meanAcc_l = colMeans(masterAcc_l)
meanSens_l = colMeans(masterSens_l)
meanSpef_l = colMeans(masterSpef_l)

k_results_l <- data.frame(kvalue = seq(1,numks,1), 
                          accuracy_avg = meanAcc_l,
                          Sensitivity_avg = meanSens_l,
                          Specificity_avg = meanSpef_l
)

head(k_results_l%>%arrange(desc(accuracy_avg)))%>%gt()%>%tab_header(title = "KNN-Leave one Out (100 iterations of K:1-30)")
```


# KNN Test/Train Cross Validation

```{r message=FALSE, warning=FALSE}

library(class)
library(caret)


attrition_table_train <- attrition_table_norm[1:696,]
attrition_table_test  <- attrition_table_norm[697:870,]

attrition_table_train_labels <- attrition_table_org[1:696,1]
attrition_table_test_labels  <- attrition_table_org[697:870,1]

attrition_table_norm_pred.6 <- knn(attrition_table_train%>%select(JobSatisfaction,MonthlyIncome,Age,EnvironmentSatisfaction,JobLevel,                                                        ,JobInvolvement,StockOptionLevel,TotalWorkingYears,YearsAtCompany,YearsInCurrentRole,YearsWithCurrManager), 
                                attrition_table_test%>%select(JobSatisfaction,MonthlyIncome,Age,EnvironmentSatisfaction,JobLevel,                                                    ,JobInvolvement,StockOptionLevel,TotalWorkingYears,YearsAtCompany,YearsInCurrentRole,YearsWithCurrManager),
                                    cl = attrition_table_train_labels$Attrition, prob = TRUE,k=6)

attrition_table_norm_pred.6 <- knn(attrition_table_train%>%select(JobSatisfaction,MonthlyIncome,Age,EnvironmentSatisfaction,JobLevel,
                                 ,JobInvolvement,StockOptionLevel,TotalWorkingYears,YearsAtCompany,YearsInCurrentRole,YearsWithCurrManager), 
                                    attrition_table_test%>%select(JobSatisfaction,MonthlyIncome,Age,EnvironmentSatisfaction,JobLevel,
                                  ,JobInvolvement,StockOptionLevel,TotalWorkingYears,YearsAtCompany,YearsInCurrentRole,YearsWithCurrManager),
                                    cl = attrition_table_train_labels$Attrition, prob = TRUE,k=6)


table(attrition_table_norm_pred.6,attrition_table_test_labels$Attrition)

confusionMatrix(table(attrition_table_norm_pred.6,attrition_table_test_labels$Attrition))

```



# Question 2

Predict the Attrition for the sample set using the best data model 


```{r message=FALSE, warning=FALSE}
no_attrition_table <- subset(no_attrition_table, select =-c(ID,EmployeeCount,Over18,StandardHours))

nCol2  = ncol(no_attrition_table)


no_attrition_table$BusinessTravel <- factor(no_attrition_table$BusinessTravel, levels =  c('Non-Travel','Travel_Rarely','Travel_Frequently'),
                                            label = c('Non-Travel','Travel_Rarely','Travel_Frequently'))

no_attrition_table$Department <- factor(no_attrition_table$Department, levels =  c('Sales','Research & Development','Human Resources'),
                                        label = c('Sales','Research & Development','Human Resources'))

no_attrition_table$EducationField <- factor(no_attrition_table$EducationField, levels =  c('Life Sciences','Medical','Technical Degree','Other','Marketing',
                                                                                           'Human Resources'),label = c('Life Sciences','Medical','Technical Degree','Other','Marketing','Human Resources'))

no_attrition_table$Gender <- factor(no_attrition_table$Gender, levels =  c('Male','Female'),label = c('Male','Female'))

no_attrition_table$JobRole <- factor(no_attrition_table$JobRole, levels =  c('Sales Executive','Sales Representative','Research Director','Research Scientist','Manufacturing Director',
                                                                             'Healthcare Representative','Manager','Human Resources','Laboratory Technician'),label = c('Sales Executive','Sales Representative','Research Director','Research Scientist',
                                                                                                                                                                        'Manufacturing Director','Healthcare Representative','Manager','Human Resources','Laboratory Technician'))
no_attrition_table$MaritalStatus <- factor(no_attrition_table$MaritalStatus,levels =  c('Single','Divorced','Married'),label = c('Single','Divorced','Married'))

no_attrition_table$OverTime <- factor(no_attrition_table$OverTime, levels =  c('No','Yes'),label = c('No','Yes'))

no_attrition_table[sapply(no_attrition_table,is.factor)]  <- data.matrix(no_attrition_table[sapply(no_attrition_table,is.factor)])

no_attrition_table_org <- no_attrition_table


data_norm <- function(x) { (x -min(x))/(max(x)-min(x))   }

no_attrition_table_norm <- as.data.frame(lapply(no_attrition_table_org[1:nCol2],data_norm))


attrition_table_train <- attrition_table_norm[1:870,]
no_attrition_table_test  <- no_attrition_table_norm[1:300,]

no_attrition_table_train_labels <- attrition_table_org[1:870,1]
no_attrition_table_test_labels  <- no_attrition_table_org[1:300,1]

no_attrition_table_norm_pred.6 <- knn(train = attrition_table_train, test = no_attrition_table_test,cl = no_attrition_table_train_labels$Attrition, k=6)

write.csv(no_attrition_table_norm_pred.6,"E://Case2PredictionsCheruvathoor_Attrition.csv", row.names = FALSE)

summary(no_attrition_table_norm_pred.6)


```

# Question 3


Predict the Salary for the 300 observations provided. 

Use the linear regression model and predict the salary based on the explonatory variables on the train data.
Identify the coeifficinets and use the model to predict the 300 observation data provided. 


```{r message=FALSE, warning=FALSE}

library(coefplot)

no_salary_table <-read_csv("E://CaseStudy2CompSetNoSalary.csv")

no_salary_table_org <- no_salary_table

no_salary_table <- subset(no_salary_table, select =-c(ID,EmployeeCount,Over18,StandardHours))


no_salary_table$BusinessTravel <- factor(no_salary_table$BusinessTravel, levels =  c('Non-Travel','Travel_Rarely','Travel_Frequently'),
                                            label = c('Non-Travel','Travel_Rarely','Travel_Frequently'))

no_salary_table$Department <- factor(no_salary_table$Department, levels =  c('Sales','Research & Development','Human Resources'),
                                        label = c('Sales','Research & Development','Human Resources'))

no_salary_table$EducationField <- factor(no_salary_table$EducationField, levels =  c('Life Sciences','Medical','Technical Degree','Other','Marketing',
                                                                                           'Human Resources'),label = c('Life Sciences','Medical','Technical Degree','Other','Marketing','Human Resources'))

no_salary_table$Gender <- factor(no_salary_table$Gender, levels =  c('Male','Female'),label = c('Male','Female'))

no_salary_table$JobRole <- factor(no_salary_table$JobRole, levels =  c('Sales Executive','Sales Representative','Research Director','Research Scientist','Manufacturing Director',
                                                                             'Healthcare Representative','Manager','Human Resources','Laboratory Technician'),label = c('Sales Executive','Sales Representative','Research Director','Research Scientist',
                                                                                                                                                                        'Manufacturing Director','Healthcare Representative','Manager','Human Resources','Laboratory Technician'))
no_salary_table$MaritalStatus <- factor(no_salary_table$MaritalStatus,levels =  c('Single','Divorced','Married'),label = c('Single','Divorced','Married'))

no_salary_table$Attrition <- factor(no_salary_table$Attrition, levels =  c('No','Yes'),label = c('No','Yes'))

no_salary_table$OverTime <- factor(no_salary_table$OverTime, levels =  c('No','Yes'),label = c('No','Yes'))

no_salary_table[sapply(no_salary_table,is.factor)]  <- data.matrix(no_salary_table[sapply(no_salary_table,is.factor)])


view(no_salary_table)

nCol3 = ncol(no_salary_table)

data_norm <- function(x) { (x -min(x))/(max(x)-min(x))   }

no_salary_table_norm <- as.data.frame(lapply(no_salary_table[1:nCol3],data_norm))

view(no_salary_table_norm)

#### Normalize the No salary data and now we are good to do the  Mnthly Income predictions

fit <- lm(MonthlyIncome ~ Age+Gender+BusinessTravel+JobRole+DailyRate+Department+DistanceFromHome+Education+
            EducationField+EmployeeNumber+EnvironmentSatisfaction+Gender+HourlyRate+JobInvolvement+
            JobLevel+JobRole+JobSatisfaction+MaritalStatus+MonthlyRate+NumCompaniesWorked+
            PercentSalaryHike+PercentSalaryHike+PerformanceRating+RelationshipSatisfaction+StockOptionLevel+
            TotalWorkingYears+TrainingTimesLastYear+WorkLifeBalance+
            YearsAtCompany+YearsInCurrentRole+YearsSinceLastPromotion+YearsWithCurrManager, data = attrition_table_norm)

summary(fit)

coefplot(fit,intercept = FALSE)

no_salary_table_norm$MonthlyIncome  = predict(fit, newdata = no_salary_table_norm)

no_salary_table_org <-  subset(no_salary_table_org, select =c(ID))

no_salary_table_org$MonthlyIncome  = as.integer(no_salary_table_norm$MonthlyIncome *  19999 )

view(no_salary_table_org)
         
write.csv(no_salary_table_org,"E://Case2PredictionsCheruvathoor_Salary.csv", row.names = FALSE)

```
